Journal

### 2024-10-24 Journal/notes.tex
Journal up to date

I consider a quadratic Lyapunov function $V = \tilde{x}\tr P \tilde{x}$ and I want to bound $\Delta V$ to be strictly negative.

>>>>

I consider a quadratic Lyapunov function $V = \tilde{x}\tr P \tilde{x}$ and I want to bound $\Delta V$ to be strictly negative.
Expressing $\tilde{x}_*^{+}$ in a compact way:

$$
  \tilde{x}^{+} = A x + B u + C \phi - A x_* - B u_* - C \phi_* = A \tilde{x} + B \tilde{u} + C \tilde{\phi}
$$

In this expression $\tilde{u}$ can be expressed with respect to $\omega$ and $x$:
$$
  \tilde{u} = N_{ux} x + N_{u \omega} + N_{ub} - N_{ux} x_* - N_{u \omega} \omega_* - N_{ub} = N_{ux} \tilde{x} + N_{u \omega} \tilde{\omega}
$$

I consider now the change of variables $\psi = \nu - \omega$ to make the problem compliant with the sector conditions used in \cite{css-extended}:

\begin{align*}
  \tilde{\nu} &= R N_{\nu x} \tilde{x} \\
  \tilde{u} &= N_{ux} \tilde{x} + N_{u \omega} (\tilde{\nu} - \tilde{\psi}) = N_{ux} \tilde{x} + N_{u \omega} R N_{\nu x} \tilde{x} - N_{u \omega} \tilde{\psi} = (N_{ux} + N_{u \omega} R N_{\nu x}) \tilde{x} - N_{u \omega} \tilde{\psi} = R_{\omega} \tilde{x} - N_{u \omega} \tilde{\psi} \\
  \tilde{x}^{+} &= (A + B R_{\omega}) \tilde{x} - B N_{u \omega} \tilde{\psi} + C \tilde{\phi} = \bar{A} \tilde{x} + \bar{B} \tilde{\psi} + C \tilde{\phi}
\end{align*}

With \begin{itemize}
  \item $R = \left( \Id - N_{\nu \omega} \right)^{-1}$
  \item $R_{\omega} = N_{ux} + N_{u \omega} R N_{\nu x}$
  \item $\bar{A} = A + B R_{\omega}$
  \item $\bar{B} = - B N_{u \omega}$
\end{itemize}

I can express then:

\begin{align*}
  \Delta V &= (\tilde{x}^{+})\tr P \tilde{x}^{+} - \tilde{x}\tr P \tilde{x} = \left( \bar{A} \tilde{x} + \bar{B} \tilde{\psi} + C \tilde{\phi} \right)\tr P \left( \bar{A} \tilde{x} + \bar{B} \tilde{\psi} + C \tilde{\phi} \right) - \tilde{x}\tr P \tilde{x} = \\
  &= \bmx{
    \tilde{x}\\
    \tilde{\psi}\\
    \tilde{\phi}
  }\tr \bmx{
    \bar{A}\tr P \bar{A} - P & \bar{A}\tr P \bar{B} & \bar{A}\tr P C \\
    \bar{B}\tr P \bar{A} & \bar{B}\tr P \bar{B} & \bar{B}\tr P C \\
    C\tr P \bar{A} & C\tr P \bar{B} & C\tr P C
  } \bmx{
    \tilde{x}\\
    \tilde{\psi}\\
    \tilde{\phi}
  }
\end{align*}

I have to include now the sector conditions to deal with the non-linearity of the activation functions of the NN controller. In this particular case we deal with only one saturation function. Following the approach in \cite{css-extended} I can express the sector condition as:

$$
  \tilde{\psi}\tr T \left[ G \tilde{x} - \tilde{\nu} + \tilde{\psi} \right] \leq 0 \to \bmx{
    \tilde{x} \\
    \tilde{\psi} \\
    \tilde{\phi}
  }\tr \bmx{
    0 & 0 & 0 & 0 \\
    TG & -T & T & 0 \\
    0 & 0 & 0 & 0
  } \bmx{
    \tilde{x} \\
    \tilde{\psi} \\
    \tilde{\nu} \\
    \tilde{\phi}
  } \to \bmx{
    \tilde{x} \\
    \tilde{\psi} \\
    \tilde{\phi}
  }\tr \bmx{
    0 & 0 & 0 & 0 \\
    TG & -T & T & 0 \\
    0 & 0 & 0 & 0
  } \bmx{
    \Id & 0 & 0 \\
    R N_{\nu x} & \Id - R & 0 \\
    0 & \Id & 0  \\
    0 & 0 & \Id
  } \bmx{
    \tilde{x} \\
    \tilde{\psi} \\
    \tilde{\phi}
  }
$$

Having written $\tilde{\nu} = N_{\nu x} \tilde{x} + N_{\nu \omega} \tilde{\omega} = N_{\nu x} \tilde{x} + N_{\nu \omega} (\tilde{\nu} - \tilde{\psi}) \to (\Id - N_{\nu \omega}) \tilde{\nu} = N_{\nu x} \tilde{x} - N_{\nu \pi} \tilde{\psi} \to \tilde{\nu} = R N_{\nu x} \tilde{x} + (\Id - R) \tilde{\psi}$ 
Exploiting the condition $N_{\nu \omega} = \Id - R^{-1} \to -R N_{\nu \omega} = \Id - R$

I can rewrite the sector conditions as:

\begin{align*}
  \bmx{
    \tilde{x} \\
    \tilde{\psi} \\
    \tilde{\phi}
  }\tr M_{1} R_{\phi} \bmx{
    \tilde{x} \\
    \tilde{\psi} \\
    \tilde{\phi}
  } \leq 0
\end{align*}

Given their sign definiteness I can subtract the term to $\Delta V$ in order to impose a negative bound to the Lyapunov function derivative:

$$
  \Delta V \leq \Delta V - He \left\{\bmx{
    \tilde{x} \\
    \tilde{\psi} \\
    \tilde{\phi}
  } M_{1} R_{\phi} \bmx{
    \tilde{x} \\
    \tilde{\psi} \\
    \tilde{\phi}
  }\right\} < 0
$$

Finally, I include the sector condition for the non-linearity $\phi = sin(\theta) - \theta$. The sector used is the following that is valid for $\theta \in \left[ -\pi, \pi \right]$:

\begin{align*}
  & \bmx{
    \tilde{\theta} \\
    sin(\tilde{\theta}) - \tilde{\theta}
  }\tr \bmx{
    0 & -1 \\
    -1 & -2
  } \bmx{
    \tilde{\theta} \\
    sin(\tilde{\theta}) - \tilde{\theta}
  } \geq 0 \to \\
  & \to \bmx{
    \tilde{x} \\
    \tilde{\psi} \\
    \tilde{\phi}
  }\tr \bmx{
    1 & 0 & 0 & 0_{1 \times n_{\phi}} & 0 \\
    & 0_{1 \times n_{x}} & & 0_{1 \times n_{\phi}} & 1
  }\tr \bmx{
    0 & -1 \\
    -1 & -2
  } \bmx{
    1 & 0 & 0 & 0_{1 \times n_{\phi}} & 0 \\
    & 0_{1 \times n_{x}} & & 0_{1 \times n_{\phi}} & 1
  } \bmx{
    \tilde{x} \\
    \tilde{\psi} \\
    \tilde{\phi}
  } > 0 \to \\
  & \to \bmx{
    \tilde{x} \\
    \tilde{\psi} \\
    \tilde{\phi}
  }\tr R_{q}\tr \bmx{
    0 & -1 \\
    -1 & -2
  } R_{q} \bmx{
    \tilde{x} \\
    \tilde{\psi} \\
    \tilde{\phi}
  } > 0
\end{align*}

I add this sector condition to $\Delta V$:

\begin{align*}
  \Delta V &\leq \Delta V - He \left\{\bmx{
    \tilde{x} \\
    \tilde{\psi} \\
    \tilde{\phi}
  } M_{1} R_{\phi} \bmx{
    \tilde{x} \\
    \tilde{\psi} \\
    \tilde{\phi}
  }\right\} + \bmx{
    \tilde{x} \\
    \tilde{\psi} \\
    \tilde{\phi}
  } R_{q}\tr \bmx{
    0 & -1 \\
    -1 & -2
  } R_{q} \bmx{
    \tilde{x} \\
    \tilde{\psi} \\
    \tilde{\phi}
  } < 0 \to \\
  & \to \bmx{
    \tilde{x} \\
    \tilde{\psi} \\
    \tilde{\phi}
  }\tr
  \left( \bmx{
    \bar{A}\tr P \bar{A} - P & \bar{A}\tr P \bar{B} & \bar{A}\tr P C \\
    \bar{B}\tr P \bar{A} & \bar{B}\tr P \bar{B} & \bar{B}\tr P C \\
    C\tr P \bar{A} & C\tr P \bar{B} & C\tr P C
  } - R_{\phi} \tr M_{1}\tr - M_{1} R_{\phi} + R_{q}\tr \bmx{
    0 & -1 \\
    -1 & -2
  } R_{q}\right)
  \bmx{
    \tilde{x} \\
    \tilde{\psi} \\
    \tilde{\phi}
  } < 0 \to \\
  & \to \bmx{
    \tilde{x} \\
    \tilde{\psi} \\
    \tilde{\phi}
  }\tr
  M
  \bmx{
    \tilde{x} \\
    \tilde{\psi} \\
    \tilde{\phi}
  } < 0 \to M < 0 
\end{align*}

As regards the ellipsoid condition I use the same as in \textit{"Linear integrator 1 dimension case"}.

\subsection*{3 layer saturation controller to non-linear pendulum with integrator in the state}

The same formulation was extended to the 3 layer network where the only things that change are the matrix $N$ and the ellipsoid conditions that now are larger: $n_{\phi}$ instead of just $1$.

After solving the LMI problem I compared the sizes of the regions of attraction (ROA) of the linear system controlled with the saturated input, the non-system controlled with the saturated input and the non-system controlled with the 3-layer NN controller.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.65\textwidth]{ROA_comparison}
    \caption{It is intuitive to think that the ROA decreases linearly with the number of constraints set to the LMI}
\end{figure}
